---
title: "ECON 390 Homework 1"
output:
  pdf_document: default
  html_notebook: default
---

## Creating Variables
1. Create a vector of submission IDs of you and your group members called “sub ids”
```{r creatingVariables1}
sub_ids = c("SID2")
sub_ids
```

2. Create a vector of the integers from 1 to 5 called “my vec”.
```{r creatingVariables2}
my_vec = c(1:5)
my_vec
```

3. Update “my vec” by adding 10 to each element of “my vec”
```{r creatingVariables3}
my_vec = my_vec + 10
my_vec
```

4. Create a vector of integers from 1994 to 2021 called “years alive.”
```{r creatingVariables4}
years_alive = c(1994:2021)
years_alive
```

5. Create a vector of logicals testing if an element in the “years alive” vector is a leap year
```{r creatingVariables5}
names(years_alive) = c(1994:2021)
ifleap = list(names(years_alive), years_alive%%4 == 0)
ifleap
```

6. Draw 1000 simulations from a N(0, 1) distribution and call it “norm draws”.
```{r creatingVariables6}
norm_draws = rnorm(1000, mean = 0, sd = 1)
```

7. Create a vector of that is the log of the draws from item 6 which is called “log draws”.
```{r creatingVariables7}
log_draws = log(norm_draws)
```

8. Create a logical vector called “draws NaN” that tests which elements of “log draws” are NaNs.
```{r creatingVariables8}
draws_NaN = is.na(log_draws)
```

9. Create a sequence of integers from 1 to N, where N is the number of NaN elements in “log draws”.
```{r creatingVariables9}
sequenceOfNaNs =  seq(which(is.na(log_draws)))
sequenceOfNaNs
```

10. Use modular arithmetic operations to create a vector called “class length” whose first element
is the number of whole hours in one of our class meetings and the second is the number of
leftover minutes. E.g. an hour and a half would be “c(1,30)”.
```{r creatingVariables10}
class_length = c(75%/%60, 75%%60)
class_length
```

## Creating Data Sets
1. Create a variable called “mtcars list” that is each variable in mtcars as a different in the list.
```{r creatingDatasets1}
mtcars_list = as.list(ls(mtcars))
mtcars_list
```

2. Create a 2 × 13 matrix of the letters of the alphabet where they are read from left to right.
Call this matrix “letter mat”.
```{r creatingDatasets2}
letter_mat = matrix(letters, nrow = 2, ncol = 13, byrow = 1)
letter_mat
```

3. Create a matrix that is 5 × 5 of new N(0, 1) draws called “norm draws mat”.
```{r creatingDatasets3}
norm_draws_mat = matrix(rnorm(n = 25, mean = 0, sd = 1), nrow = 5, ncol = 5)
norm_draws_mat
```

4. Using the “iris” data set, store the Species variable as a factor variable named “iris species”
where “versicolor” is the first level, “virginica” is the second, and “setosa” is the third.
```{r creatingDatasets4}
iris_species = factor(iris$Species, levels = c("versicolor", "virginica", "setosa"))
iris_species
```

5. Create a sequence of numbers from 1.2 to 5.3 by 0.05 called “first seq”.
```{r creatingDatasets5}
first_seq = seq(from = 1.2, to = 5.3, by = 0.05)
first_seq
```

6. Create a sequence of numbers from 1.2 to 5.3 that is 100 elements long called “second seq”.
```{r creatingDatasets6}
second_seq = seq(from = 1.2, to = 5.3, length.out = 100)
second_seq
```

7. Load the AER library and attach the GSOEP9402 data set (data(“GSOEP9402”)). What are
the dimensions of the data set? What are the names of the variables? Rename the variables
to the same name, but all in uppercase rather than lowercase.
```{r creatingDatasets7}
library(AER)
data("GSOEP9402")
dim(GSOEP9402)
toupper(names(GSOEP9402))
```

8. Create the following vectors:
• “person id” which is a numeric vector from 1 to 50 that increases by 1.
• “time” which is a numeric vector from 2001 to 2020 that increases by 1.
```{r creatingDatasets8}
person_id = seq(from = 1, to = 50, by = 1)
person_id
time = seq(from = 2001, to = 2020, by = 1)
time
```

9. Finally, create a data.frame called “panel data” that have the following variables:
• “id” where each element of “person id” is repeated 20 times.
• “time” where the time vector is repeated 50 times.
• “draw” that is the vector “norm draws”.
• Note: this data set is pointless; this is just to get you familiar with creating data.frames
```{r creatingDatasets9}
time = rep(time, times=50)
id = rep(person_id, times  = 20)
panel_data = data.frame(id, time, norm_draws)
```

## Indexing

1. Store a vector of the id’s of the elements in“draws NaN” that are NaN called “NaN ids”.
```{r indexing1}
nan_Ids = which(draws_NaN == T)
```

2. Store a vector of the id’s of the elements in “norm draws” that are positive called “pos ids”.
```{r indexing2}
pos_ids = which(norm_draws > 0)
```

3. Store the elements with even indices from “norm draws” called “even id draws”.
```{r indexing3}
even_id_draws = which(norm_draws %% 2 == 0)
```

4. Extract the value from the 20th observation of the “mpg” variable in the “mtcars” data set
using three different ways we discussed in class.
```{r indexing4}
mtcars$mpg[20]
mtcars[[1]][20]
mtcars[["mpg"]][20]
```

5. Extract the value from the 17th observation of the “cyl” and “hp” variables in the “mtcars”
data set.
```{r indexing5}
c(mtcars$cyl[17], mtcars$hp[17])
```

6. (Harder) Extract all observations from “mtcars” that have 8 cylinders (cyl) using logical
expressions/indexing.
```{r indexing6}
subset(mtcars, mtcars$cyl == 8)
```

7. Load the “Titanic” data set. This data set documents the survivors and victims of The
Titanic. Use the “str()” function to determine the dimensions and structure of the data set.
Print the outcomes for the following subgroups of passengers:
(a) The outcomes of the female, adult, crew.
(b) The outcomes for the male, first-class, adults.
(c) The outcomes of the female, first-class, children.
(d) The outcomes of the male, second-class, adults.
```{r indexing7}
str(Titanic)
Titanic = data.frame(Titanic)
partA = Titanic[Titanic$Sex== "Female" & Titanic$Age== "Adult" & Titanic$Class =="Crew",]
partB = Titanic[Titanic$Sex== "Male" & Titanic$Class =="1st" & Titanic$Age== "Adult",]
partC = Titanic[Titanic$Sex== "Female" & Titanic$Class =="1st" & Titanic$Age== "Child",]
partD = Titanic[Titanic$Sex== "Male" & Titanic$Class =="2nd" & Titanic$Age== "Adult",]
```

8. Using “mtcars list”, extract the “wt” as an atomic vector using two different ways of indexing
a list discussed in class.
```{r indexing8}
mtcars[[mtcars_list[[11]]]]
```

9. Using “mtcars list”, extract the “wt” variable as a list.
```{r indexing9}
mtcars[mtcars_list[[11]]]
```